Data Analysis Methods for the Detection of the Epoch of Reionization

### 2017-11-15 PAPER.tex
made more edits from Danny up until Section 3.3

We now present a case study of these same three themes using data from the PAPER experiment. We use the intuition we've developed through our toy model simulations in order to make educated data analysis choices for PAPER. In light of our new understandings, we also highlight major changes since \citet{ali_et_al2015} which have contributed to a revised power spectrum upper limit result.

As a brief review, PAPER is a dedicated $21$ cm experiment located in the Karoo Desert in South Africa. The PAPER-64 configuration consists of 64 dual-polarization drift-scan elements (Figure \ref{fig:paper}) that are arranged in a grid layout. For our case study, we focus solely on Stokes I data from PAPER's $30$ m East/West baselines. For information about the backend system of PAPER-64, its observations, and data reduction pipeline, we refer the reader to \citet{parsons_et_al2010} and \citet{ali_et_al2015}.

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 0.3cm 0.3cm 0.3cm},width=\columnwidth]{plots/paper_dipole.png}
	\caption{PAPER dipole in South Africa.}
	\label{fig:paper}
\end{figure}

@@

\noindent of $(22.4$ mK$)^{2}$ in the range $0.15 < k < 0.5$ h Mpc$^{-1}$ at $z = 8.4$. The revision of this limit stems mostly from previously underestimated signal loss and underestimated error bars \cc{cite retraction paper}, both of which we address in the following sections.

For our analysis, we use $8$ hours of LST (RA $0.5$-$8.6$ hours) and $51$ total baselines (\citet{ali_et_al2015} uses an RA range of $0$-$8.6$ hours). All power spectrum results are produced using channels $95$-$115$, corresponding to a center redshift of $z=8.4$, identical to the analysis in \citet{ali_et_al2015}. We note that the PAPER-64 dataset that we use in this case study differs from that in \citet{ali_et_al2015} mainly by the fringe-rate filter. In \citet{ali_et_al2015}, the applied filter was degraded by widening it in fringe-rate space. This was chosen in order to increase the number of independent modes and reduce signal loss. With the development of a robust method for assessing signal loss, we feel comfortable using a narrow filter --- the optimal fringe-rate filter --- in order to maximize sensitivity. This filter is computed for a fiducial $30$ m baseline at $150$ MHz, the center frequency in our band. The filter in both the fringe-rate domain and time domain is shown in Figure \ref{fig:frp}.

@@

\subsection{Case Study: Signal Loss}
\label{sec:Sigloss}

In Section \ref{sec:SiglossOverview}, we showed how signal loss arises when weighting data using information from itself. Here we describe a methodology that simulates the injection and recovery of a cosmological signal in order to quantify the amount of signal loss accompanying a weighting scheme. In particular, we highlight major differences from the signal loss computation used in \citet{ali_et_al2015}, which previously underestimated losses.

@@

We simulate an EoR signal by creating a random Gaussian signal (with a default variance of $1$) with the same shape as our data, and we fringe-rate filter this signal twice using the optimal filter. The first filter transforms the white noise into a signal that's attached to the sky (i.e. what our instrument observes). The second filter represents the fringe-rate filtering step in our data analysis pipeline. This mock EoR signal is injected on top of fringe-rate filtered PAPER-64 data at a range of amplitude levels that spans the data level.

@@

The relationship between the input and output power spectra, $P_{in}$ and $P_{out}$, can be thought of as a transfer function that maps an output power spectrum distribution into an input distribution. By viewing data through this signal loss lens, we are asking the question ``what input power spectrum distribution could this (signal-loss affected) data come from?" In the next section, we describe our signal loss transfer function for PAPER-64 data and detail the computation used to translate our power spectrum result into one viewed through a signal loss lens.

@@

We now shift our attention towards computing signal loss for our fringe-rate filtered PAPER-64 dataset. While our methodology outlined below is robust to any weighting scheme, here we demonstrate the computation using inverse covariance weighting ($\textbf{R} = \textbf{C}^{-1}$), the weighting scheme used in \citet{ali_et_al2015} which leads to substantial loss. With this weighting, our expressions for $P_{in}$ and $P_{out}$ become:

@@

One peculiar aspect of Figure \ref{fig:sigloss_transfercurve} is the fact that at low $P_{in}$ values it appears that we can have signal gain ($P_{out} > P_{in}$). This is unphysical in nature but caused due to the cross-terms dominating in Equation \eqref{eq:pout_expand} once $\textbf{e}$ becomes small. As $\textbf{e}$ increases, we move into a regime where $P_{out} \sim P_{in}$, and then eventually into a regime where $P_{out} < P_{in}$ when $\textbf{e}$ is large enough to destroyed if weighting the data using itself. Although we only show figures for one $k$ value, we note that the shape of the transfer curve is nearly identical for all $k$'s (though we treat each $k$ separately).

@@

To summarize this section, we also show power spectrum results for fringe-rate filtered PAPER-64 data before and after signal loss correction in Figure \ref{fig:ps2_data}, using inverse covariance weighting. The blue dashed line represents the unweighted power spectrum, which is identical in both panels (an important check, as we expect no signal loss for this case). The black and grey points are positive and negative power spectrum values plotted with $2\sigma$ error bars. Prior to signal loss correction, it is clear that the power spectrum is unfeasible because it is well below the theoretical noise level prediction (solid green curve). Post-correction, the power spectrum values blow up to be higher than both the thEoRy and unweighted power spectrum.

@@

We saw in Section \ref{sec:otherweight} how limiting the number of down-weighted eigenmodes (i.e. flattening out part of the eigenspectrum) can help minimize signal loss. We experiment with this idea on PAPER-64 data, dialing the number of modes down-weighted from zero (which is equivalent to identity-weighting, or the unweighted case) to $21$ (which is full inverse covariance weighting of our $21$ channels). The power spectrum results for one $k$ value, both before and after signal loss correction, are shown in Figure \ref{fig:sigloss_modeloop}. We see that the amount of signal loss increases as weighting becomes more aggressive (grey curve). In other words, more random fluctuations are being overfit and subtracted as more modes are down-weighted. We also find that the power spectrum upper limit, post signal loss correction, increases with the number of down-weighted modes (black curve). As discussed previously, we are limited in our power spectrum accuracy by the number of independent samples in our dataset and our errors dramatically increase as we surpass the $\sim8$ samples that we have in fringe-rate filtered PAPER-64 data.

Optimistically, we expect there to be a `sweet spot' as we dial our regularization knob; a level of regularization where weighting is beneficial compared to not weighting (blue dashed line). We see in Figure \ref{fig:sigloss_modeloop} that this occurs when only the strongest $\sim3-4$ eigenmodes are down-weighted, though the improvement from the unweighted case is not significant.

@@

In this section we have shown three simple ways of regularizing $\hat{\textbf{C}}$ to minimize signal loss using PAPER-64 data. There are many other weighting schemes that we leave for consideration in future work. For example, one could estimate $\hat{\textbf{C}}$ using information from different subsets of baselines. For redundant arrays this could mean calculating $\hat{\textbf{C}}$ from a different but similar baseline type, such as the $\sim30$m diagonal PAPER baselines (instead of the horizontal E/W ones). Alternately, covariances could be estimated from all other baselines except the two being cross-multiplied when forming a power spectrum estimate. This method was used in \citet{parsons_et_al2014} in order to avoid suppressing the $21$cm signal, and it's worth noting that the PAPER-32 results are likely safe from the issue of signal loss underestimation, unlike PAPER-64.

Another possible way to regularize $\hat{\textbf{C}}$ is to use different ranges of LST. For example, one could calculate $\hat{\textbf{C}}$ with data from LSTs where foregrounds are stronger (earlier or later LSTs than the `foreground-quiet' range used in forming power spectra) --- doing so may yield a better description of the foregrounds that we desire to down-weight. Fundamentally, each of these examples are similar in that they rely on a computation of $\hat{\textbf{C}}$ from data that is similar but not exactly the same as the data that is being down-weighted. Ideally this would be effective in down-weighting shared contaminants yet avoid signal loss from over-fitting modes in the power spectrum dataset itself.

@@

\subsection{Case Study: Error Estimation}
\label{sec:Error}

@@

Theoretical errors of noise, computed from the instrumental temperature and observational parameters, provide a useful cross-check for bootstrapped errors and is a helpful predictor of the maximum sensitivity that can be achieved by a particular observation and analysis. Here we walk through a detailed computation of a theoretical noise estimation as applied to PAPER-64 observations and highlight major changes from \citet{ali_et_al2015}. We first give a general formula, and then go into detail about each factor.

The sensitivity prediction for a power spectral analysis of interferometric $21$ cm data, in temperature-units, is:

@@

\begin{itemize}
\item $X^{2}Y$: Conversion factors from observing coordinates (angles on the sky) to cosmological coordinates (co-moving distances). For $z=8.4$, $X^{2}Y = 5 \times 10^{11} h^{-3} Mpc^{3} str^{-1} GHz^{-1}$.
\item $\Omega_{eff}$: The primary beam angular size. The effective beam area changes with the application of a fringe-rate filter, since parts of the beam are up-weighted and down-weighted. Using numbers from Table 1 in \citet{parsons_et_al2016}, $\Omega_{eff} = 0.74^{2}/0.24$ for an optimal fringe-rate filter. 
\item $T_{sys}$: The system temperature is set by:

@@

where $\nu$ are frequencies in GHz. We use a receiver temperature of $200$ K, yielding $T_{sys} = 487$ at $150$ MHz. This is lower than in \citet{ali_et_al2015} because \cc{why?}.
\item $\sqrt{2}$: This factor in the denominator of the sensitivity equation comes from taking the real part of the power spectrum estimates after squaring visibility measurements. In \citet{ali_et_al2015}, a factor of $2$ was mistakenly used, which would imply that that real parts were obtained prior to squaring.
\item $N_{lsts}$: The number of LST hours that go into a power spectrum estimation. The sensitivity scales as the square root because we integrate incoherently over time. For PAPER-64, $N_{lsts} = 8$ hours.
\item $N_{seps}$: The number of baseline separation types averaged incoherently in a final power spectrum estimate. For the analysis in this paper, we only use one type of baseline, hence $N_{seps}=1$. The revised limit in \cc{cite kolopanis paper} uses three separation types.
\item $t_{int}$: The integration time of the data. It is crucial to adapt this number if filtering is applied along the time axis (i.e. a fringe-rate filter). We compute the effective integration time of our fringe-rate filtered data by scaling the original integration time using the following:
\begin{equation}
t_{frf} = t_{int} \frac{\int1 \, df}{\int w^{2}(f) \,df},
\end{equation}
where $t_{int}=43$ seconds, $t_{frf}$ is the fringe-rate filtered integration time, $w$ is the fringe-rate profile, and the integral is taken over all fringe-rates. In other words, we calculate the effective integration time by computing how our beam is altered in fringe-rate space. For PAPER-64, this number is $t_{int} = 3857$ s. 
\item $N_{days}$: The total number of days of data analyzed. In \citet{ali_et_al2015}, this number was set to $135$. However, because we divide our data in half (to form `even' and `odd' datasets), this number should be reduced by a factor of $2$. Additionally, because our $LST$ coverage is not $100\%$ complete (it doesn't overlap for every single day), we compute a realistic value of $N_{days}$ as the RMS of all the daily counts in one dataset (`even', for example). For PAPER-64, our revised estimate of $N_{days}$ is $\sim34$ days.
\item $N_{bls}$: The number of baselines contributing to the sensitivity of a power spectrum estimate. In \citet{ali_et_al2015}, this number was the total number of $30$ m East/West baselines used in the analysis. However, using the total number of baselines neglects the fact that we divide baselines into $5$ groups before cross-multiplying data. Our revised estimate for the parameter is $N_{bls} = \frac{N_{bls}}{N_{gps}}\sqrt{N_{gps}^{2}-N_{gps}}$, where $N_{gps} = 5$. This expression arises because $N_{bls}$ must be scaled by the total number of cross-multiplications going into the power spectrum estimate. For $5$ groups, there are $25$ total cross-products, minus $5$ cross-products since we avoid multiplying the same group with itself. For our PAPER-64 analysis with only one baseline separation type, this becomes $N_{bls} \sim 46$. 
\item $N_{pols}$: The number of polarizations averaged together. For the case of Stokes I, $N_{pols}=2$.
\end{itemize}

Equation \eqref{eq:sense} represents the most sensitive estimate for $\hat{\textbf{p}}(k)$ given our observing parameters. An additional factor of $\sqrt{2}$ is gained in sensitivity when folding our power spectra into $\Delta^{2}(k)$, due to averaging together positive and negative $k$'s.

@@

To verify our thermal noise prediction, we form power spectra estimates using a pure noise simulation. We create Gaussian random noise and use $T_{sys}$ in order to best represent the thermal noise level of PAPER-64 data. We convert $T_{sys}$ into a variance statistic using:

@@

\noindent where $\Delta\nu$ is channel spacing, $\Delta t$ is integration time, $N_{days}$ is the number of daily counts for a particular time and frequency that went into our LST binned set, and $N_{pols}$ is the number of polarizations ($2$ for Stokes I). This rms temperature sets the variance of the Gaussian random noise.

We fringe-rate filter the noise in the same way as PAPER-64 data, using an optimal fringe-rate filter. Power spectrum results for the noise simulation are shown in Figure \ref{fig:ps_noise}, where the black and grey points represent positive and negative power spectrum values, respectively (with $2\sigma$ error bars and weighting matrix $\hat{\textbf{C}}_{eff}$), the dashed blue line represents the unweighted power spectrum, and the solid green line denotes our $2\sigma$ theoretical noise prediction as calculated by Equation \eqref{eq:sense}. All three are in agreement, validating our analytical thermal noise calculation.

@@

As discussed in Section \ref{sec:ErrorOverview}, bootstrapping is only a valid way of estimating errors if a dataset is comprised of independent samples. The PAPER-64 pipeline outputs $20$ bootstraps (over baselines), each a $2$-dimensional power spectrum that is a function of $k$ and time. In \citet{ali_et_al2015}, a second round of bootstrapping occurs over the baseline bootstrap and time axes simultaneously. Random values are sampled with replacement along both axes, drawing as many values as there are number of bootstraps and times. Final power spectrum limits are then computed by taking the mean and standard deviation over this second bootstrap axis (which is typically large, on the order of $N_{boot} \sim 400$).

Because our dataset is fringe-rate filtered, we have substantially increased our integration time, resulting in fewer independent samples (from $\sim700$ samples to $8$ samples). A random draw of $700$ samples from this dataset therefore has many repeated samples, and the variance between hundreds ($N_{boot}$) of these random samples is smaller than the true underlying variance of the data. Another way of describing this underestimation is that there are fewer ways to randomly sample data with $8$ independent modes versus $700$, so bootstrapping a fringe-rate filtered dataset yields a narrower distribution of values from which we calculate its standard deviation.

To prevent error underestimation, we simply take the average along the time axis in our revised pipeline rather than bootstrapping. Hence, our final power spectrum errors are calculated from the standard deviation over the baseline bootstrapping axis only, which is still a valid axis since each baseline makes an independent measurement of the sky.

The second aspect of bootstrapping discussed in Section \ref{sec:ErrorOverview} is the overestimation of errors which arises from repeated values in a randomly sampled bootstrap. In order to maximize sensitivity, we have revised the method by which we sample baselines when bootstrapping to ensure mostly independent samples.

First, we speed things up in our pipeline by separating our total number of baselines into $5$ groups (yielding $10$ baselines per group), where there are no repeated baselines within or between groups. In \citet{ali_et_al2015}, each group is then sampled with replacement to create a new group of the same size, which can have repeated baselines inside it. In doing so, we are sacrificing some of our sensitivity since there ends up being $3$-$4$ repeated baselines per group (Figure \ref{fig:toy_error2} shows the fraction of independent samples to be $\sim55\%$ for $10$ total independent samples) . In order to maximize our sensitivity but still apply random sampling for use in error estimation, we instead form new groups using all independent baselines except the very last one, which we fill randomly. We use this method for every bootstrap, creating new groups and sampling the last spot randomly each time.

Power spectrum estimates showcasing these bootstrapping changes for PAPER-64 fringe-rate filtered data are shown in Figure \ref{fig:data_errors}. The estimates are unweighted in order to disentangle the effects of bootstrapping from signal loss. We show upper limits ($2\sigma$ error bars) for three different results. Our revised error estimation method is in black, where we only bootstrap along the baseline axis and we maximize sensitivity by random sampling only the last baseline in each group. The blue curve uses a similar method except all baselines are randomly sampled, leading to a higher power spectrum estimate because some baselines are repeated per group. In red, we bootstrap along both the baseline and time axes, drawing as many samples as there are in the data. We under-estimate errors in doing so because we are drawing many more samples than independent ones.

@@

\subsection{Case Study: Bias}
\label{sec:Bias}

@@

Although current $21$cm published power spectrum upper limits lie a $\sim$couple orders of magnitude above predicted EoR levels, ongoing analyses of deeper sensitivity datasets from PAPER, MWA, and LOFAR, as well as next generation instruments like HERA, expect to continue to push towards EoR sensitivities. As the field progresses towards a coveted detection, we have shown that it is crucial for future analyses to have a rigorous understanding of signal loss in an analysis pipeline, be able to accurately and robustly calculate both power spectrum and theoretical errors, and consistently undergo a comprehensive set of jackknive and null tests.

In particular, in this paper we have investigated the subtleties and tradeoffs of common $21$cm power spectrum techniques on signal loss, error estimation, and bias, which can be summarized as follows:

\begin{itemize}
\item Substantial signal loss can result when weighting data using empirically estimated covariances. Loss of the $21$cm signal is especially significant if more eigenmodes are down-weighted than exist in the data (having an under-constrained covariance matrix). Hence, there exists a trade-off between sensitivity driven time-averaging techniques such as fringe-rate filtering and signal loss. 
\item Signal injection and recovery simulations can be used to quantify signal loss. However, a signal-only simulation can underestimate loss by failing to account for spurious correlations between the data and signal. 
\item Errors that are estimated via bootstrapping can be underestimated if samples in the dataset are correlated. If the number of independent samples in a dataset is well-determined, bootstrapping is a simple and accurate way of estimating errors.
\item An EoR detection must pass all null tests. Similarly, performing jackknife tests along multiple axes of a dataset is necessary for confidence in a true EoR detection, and can also be used to tease out systematics.
\end{itemize}

>>>>

We now present a case study of these same three themes using data from the PAPER experiment. We use the intuition developed through the toy model simulations in order to make a new analysis of the PAPER-64 dataset originally presented in \citet{ali_et_al2015} and obtain a revised power spectrum estimate.

As a brief review, PAPER is a dedicated $21$ cm experiment located in the Karoo Desert in South Africa. The PAPER-64 configuration consists of 64 dual-polarization drift-scan elements that are arranged in a grid layout. For our case study, we focus solely on Stokes I data from PAPER's $30$ m East/West baselines. For information about the backend system of PAPER-64, its observations, and data reduction pipeline, we refer the reader to \citet{parsons_et_al2010} and \citet{ali_et_al2015}.

@@

\noindent of $(22.4$ mK$)^{2}$ in the range $0.15 < k < 0.5$ h Mpc$^{-1}$ at $z = 8.4$. %The revision of this limit stems mostly from previously underestimated signal loss and underestimated error bars \cc{cite retraction paper}, both of which we address in the following sections.

For our analysis, we use $8$ hours of LST (RA $0.5$-$8.6$ hours) and $51$ total baselines (\citet{ali_et_al2015} uses an RA range of $0$-$8.6$ hours). All power spectrum results are produced for a center frequency of $151$ MHz using a width of $10$ MHz ($20$ channels), identical to the analysis in \citet{ali_et_al2015}. We note that the PAPER-64 dataset that we use in this case study differs from that in \citet{ali_et_al2015} mainly by the fringe-rate filter. \cc{Danny says the next part is wrong... } In \citet{ali_et_al2015}, the applied filter was degraded by widening it in fringe-rate space. This was chosen in order to increase the number of independent modes and reduce signal loss. With the development of a robust method for assessing signal loss, we feel comfortable using a narrow filter --- the optimal fringe-rate filter --- in order to maximize sensitivity. This filter is computed for a fiducial $30$ m baseline at $150$ MHz, the center frequency in our band. The filter in both the fringe-rate domain and time domain is shown in Figure \ref{fig:frp}.

@@

\subsection{PAPER-64: Signal Loss}
\label{sec:Sigloss}

In Section \ref{sec:SiglossOverview}, we showed how signal loss arises when weighting data using information from the data itself. Here we describe a methodology that simulates the injection and recovery of a cosmological signal in order to quantify the amount of signal loss accompanying a weighting scheme. In particular, we highlight major differences from the signal loss computation used in \citet{ali_et_al2015}, which previously underestimated losses.

@@

To capture the full statistical likelihood of signal loss, one requires a quick way to generate many realizations of simulated $21$ cm signal visibilities. Here we use the same method used in \citet{ali_et_al2015}, where gaussian noise visibilities are filtered in fringe-rate space to retain only ``sky-like" time-modes. This signal is then added to the data.  \footnote{One specific change from \citet{ali_et_al2015} is that we add this simulated signal into the chain before the final fringe-rate filter is applied to the data. Previously, the addition was done after that final fringe-rate filter step.  This change results in an increased estimate of signal loss, likely due to the use of the fringe-rate filter as a simulator. Whether the loss would be estimated at the same level if different $21$ cm simulations were used remains an open question.}

@@

The relationship between the input and output power spectra, $P_{in}$ and $P_{out}$, can be thought of as a transfer function which, for a sampling of $P_{in}$ and $P_{out}$ provides a mapping from an output power spectrum distribution into an input distribution. By viewing data through this signal loss lens, we are asking the question ``what input power spectrum distribution could this (signal-loss affected) data come from?" In the next section, we describe our signal loss transfer function for PAPER-64 data and detail the computation used to translate our power spectrum result into one viewed through a signal loss lens.

@@

We now shift our attention towards computing signal loss for the fringe-rate filtered PAPER-64 dataset. While our methodology outlined below is robust to any weighting scheme, here we demonstrate the computation using inverse covariance weighting ($\textbf{R} = \textbf{C}^{-1}$), the weighting scheme used in \citet{ali_et_al2015} which leads to substantial loss. With this weighting, our expressions for $P_{in}$ and $P_{out}$ become:

@@

One peculiar aspect of Figure \ref{fig:sigloss_transfercurve} is the fact that at low $P_{in}$ values it appears that we can have signal gain ($P_{out} > P_{in}$). This is unphysical in nature but caused due to the cross-terms dominating in Equation \eqref{eq:pout_expand} once $\textbf{e}$ becomes small. As $\textbf{e}$ increases, we move into a regime where $P_{out} \sim P_{in}$, and then eventually into a regime where $P_{out} < P_{in}$ when $\textbf{e}$ is large enough to be destroyed if weighting the data using itself. Although we only show figures for one $k$ value, we note that the shape of the transfer curve is nearly identical for all $k$'s (though we treat each $k$ separately).

@@

To summarize this section, we also show power spectrum results for fringe-rate filtered PAPER-64 data before and after signal loss correction in Figure \ref{fig:ps2_data}, using inverse covariance weighting. The blue dashed line represents the unweighted power spectrum, which is identical in both panels (an important check, as we expect no signal loss for this case). The black and grey points are positive and negative power spectrum values plotted with $2\sigma$ error bars. Prior to signal loss correction, it is clear that the power spectrum is unfeasible because it is well below the theoretical noise level prediction (solid green curve). Post-correction, the power spectrum values blow up to be higher than both the theory and unweighted power spectrum.

@@

We saw in Section \ref{sec:otherweight} how limiting the number of down-weighted eigenmodes (i.e. flattening out part of the eigenspectrum) can help minimize signal loss. We experiment with this idea on PAPER-64 data, dialing the number of modes down-weighted from zero (which is equivalent to identity-weighting, or the unweighted case) to $21$ (which is full inverse covariance weighting of our $21$ channels). The power spectrum results for one $k$ value, both before and after signal loss correction, are shown in Figure \ref{fig:sigloss_modeloop}. We see that the amount of signal loss increases as weighting becomes more aggressive (grey curve). In other words, more random fluctuations are being overfit and subtracted as more modes are down-weighted. We also find that the power spectrum upper limit, post signal loss correction, increases with the number of down-weighted modes (black curve). As discussed previously, we are limited in our power spectrum accuracy by the number of independent samples in our dataset and our errors dramatically increase as we surpass the $\sim8$ independent samples that we have in fringe-rate filtered PAPER-64 data.

Optimistically, we expect there to be a `sweet spot' as we dial our regularization knob; a level of regularization where weighting is beneficial compared to not weighting (blue dashed line). We see in Figure \ref{fig:sigloss_modeloop} that this occurs when only the strongest $\sim3$-$4$ eigenmodes are down-weighted, though the improvement from the unweighted case is not significant.

@@

In this section we have shown three simple ways of regularizing $\hat{\textbf{C}}$ to minimize signal loss using PAPER-64 data. There are many other weighting schemes that we leave for consideration in future work. For example, one could estimate $\hat{\textbf{C}}$ using information from different subsets of baselines. For redundant arrays this could mean calculating $\hat{\textbf{C}}$ from a different but similar baseline type, such as the $\sim30$ m diagonal PAPER baselines (instead of the horizontal E/W ones). Alternately, covariances could be estimated from all other baselines except the two being cross-multiplied when forming a power spectrum estimate. This method was used in \citet{parsons_et_al2014} in order to avoid suppressing the $21$ cm signal, and it's worth noting that the PAPER-32 results are likely safe from the issue of signal loss underestimation because of this very reason.

Another possible way to regularize $\hat{\textbf{C}}$ is to use information from different ranges of LST. For example, one could calculate $\hat{\textbf{C}}$ with data from LSTs where foregrounds are stronger (earlier or later LSTs than the `foreground-quiet' range used in forming power spectra) --- doing so may yield a better description of the foregrounds that we desire to down-weight. Fundamentally, each of these examples are similar in that they rely on a computation of $\hat{\textbf{C}}$ from data that is similar but not exactly the same as the data that is being down-weighted. Ideally this would be effective in down-weighting shared contaminants yet avoid signal loss from over-fitting modes in the power spectrum dataset itself.

@@

\subsection{PAPER-64: Error Estimation}
\label{sec:Error}

@@

Re-analysis of the PAPER-64 data included a detailed study using several independently generated noise simulations. What we found was that these simulations all agreed but were discrepant with the previous analytic sensitivity calculations. The analytic calculation is only an approximation, however the differences were large enough (factors of $10$ in some cases) to warrant a careful investigation. The analytic calculation attempts to combine a large number of pieces of information in an approximate way, and when re-considering some of the approximations we have found there to be large effects. What follows here is an accounting of the differences which have been discovered.

The sensitivity prediction \cc{need refs} for a power spectral analysis of interferometric $21$ cm data, in temperature-units, is:

@@

\begin{itemize}
\item $X^{2}Y$: Conversion factors from observing coordinates (angles on the sky) to cosmological coordinates (co-moving distances). For $z=8.4$, $X^{2}Y = 5 \times 10^{11} h^{-3} Mpc^{3} str^{-1} GHz^{-1}$.
\item $\Omega_{eff}$: The effective primary beam area in steradians \footnote{\cc{ref equation}}. The effective beam area changes with the application of a fringe-rate filter, since parts of the beam are up-weighted and down-weighted. Using numbers from Table 1 in \citet{parsons_et_al2016}, $\Omega_{eff} = 0.74^{2}/0.24$ for an optimal fringe-rate filter. 
\item $T_{sys}$: The system temperature is set by:

@@

where $\nu$ are frequencies in GHz. We use a receiver temperature of $200$ K, yielding $T_{sys} = 487$ K at $150$ MHz. This is lower than in \citet{ali_et_al2015} because of previous failure to correct for using a median statistic.
\item $\sqrt{2}$: This factor in the denominator of the sensitivity equation comes from taking the real part of the power spectrum estimates after squaring visibility measurements. In \citet{ali_et_al2015}, a factor of $2$ was mistakenly used.
\item $N_{lsts}$: The number of LST hours that go into a power spectrum estimation. The sensitivity scales as the square root because we integrate incoherently over time. For PAPER-64, $N_{lsts} = 8$ hours.
\item $N_{seps}$: The number of baseline separation types averaged incoherently in a final power spectrum estimate. For the analysis in this paper, we only use one type of baseline, hence $N_{seps}=1$. The revised limit in \cc{cite kolopanis paper} uses three separation types.
\item $t_{int}$: The integration time of the data. It is crucial to adapt this number if filtering is applied along the time axis (i.e. a fringe-rate filter). We compute the effective integration time of our fringe-rate filtered data by scaling the original integration time using the following:
\begin{equation}
t_{frf} = t_{int} \frac{\int1 \, df}{\int w^{2}(f) \,df},
\end{equation}
where $t_{int}=43$ seconds, $t_{frf}$ is the fringe-rate filtered integration time, $w$ is the fringe-rate profile, and the integral is taken over all fringe-rates. For PAPER-64, this number is $t_{int} = 3857$ s. 
\item $N_{days}$: The total number of days of data analyzed. In \citet{ali_et_al2015}, this number was set to $135$. However, because we divide our data in half (to form `even' and `odd' datasets), this number should be reduced by a factor of $2$. Additionally, because our LST coverage is not $100\%$ complete (it doesn't overlap for every single day), we compute a realistic value of $N_{days}$ as:
\begin{equation}
 \frac{1}{N_{days}} = \sqrt{\Big\langle\frac{1}{N_{i}^{2}} \Big\rangle_{i}},
 \end{equation}
\noindent where $i$ is over LST (\citealt{jacobs_et_al2015}). For PAPER-64, our revised estimate of $N_{days}$ is $\sim34$ days.
\item $N_{bls}$: The number of baselines contributing to the sensitivity of a power spectrum estimate. In \citet{ali_et_al2015}, this number was the total number of $30$ m East/West baselines used in the analysis. However, using the total number of baselines neglects the fact that baselines are averaged into groups before cross-multiplying data. Our revised estimate for the parameter is $N_{bls} = \frac{N_{bls}}{N_{gps}}\sqrt{N_{gps}^{2}-N_{gps}}$, where, for the \citet{ali_et_al2015} analysis, $N_{gps} = 5$. Each baseline group averages down linearly as the number of baselines entering the group ($N_{bls}/N_{gps}$) and then as square root of the number of cross multiplied pairs ($\sqrt{N_{gps}^{2} - N_{gps}}$). For the revised PAPER-64 analysis with only one baseline separation type, this becomes $N_{bls} \sim 46$. 
\item $N_{pols}$: The number of polarizations averaged together. For the case of Stokes I, $N_{pols}=2$.
\end{itemize}

An additional factor of $\sqrt{2}$ is gained in sensitivity when folding our power spectra into $\Delta^{2}(k)$, due to averaging together positive and negative $k$'s.

@@

To verify our thermal noise prediction, we form power spectra estimates using a pure noise simulation. We create Gaussian random noise assuming a constant $T_{sys}$ (neglecting sky variation) but accounting for the true $N_{days}$ as determined by LST sampling counts for each time and frequency in the LST-binned data. We convert $T_{sys}$ into a variance statistic using:

@@

\noindent where $\Delta\nu$ is channel spacing, $\Delta t$ is integration time, $N_{days}$ is the number of daily counts for a particular time and frequency that went into our LST binned set, and $N_{pols}$ is the number of polarizations ($2$ for Stokes I). This RMS temperature sets the variance of the Gaussian random noise.

Power spectrum results for the noise simulation are shown in Figure \ref{fig:ps_noise}, where the black and grey points represent positive and negative power spectrum values, respectively (with $2\sigma$ error bars and weighting matrix $\hat{\textbf{C}}_{eff}$), the dashed blue line represents the unweighted power spectrum, and the solid green line denotes our $2\sigma$ theoretical noise prediction as calculated by Equation \eqref{eq:sense}. All three are in agreement, validating our analytical thermal noise calculation.

@@

As discussed in Section \ref{sec:ErrorOverview}, bootstrapping is only a valid way of estimating errors if a dataset is comprised of independent samples. The PAPER-64 pipeline outputs $20$ bootstraps (over baselines), each a $2$-dimensional power spectrum that is a function of $k$ and time. In \citet{ali_et_al2015}, a second round of bootstrapping occurred over the baseline bootstrap and time axes simultaneously. Random values were sampled with replacement along both axes, drawing as many values as there are number of time integrations. Final power spectrum limits were then computed by taking the mean and standard deviation over this second bootstrap axis (which is typically large, on the order of $N_{boot} \sim 400$). We emphasize that in this previous version, the number of elements sampled per bootstrap (number of times $\sim$ $700$) greatly exceeds the number of independent LST samples, under-estimating errors.

Because our dataset is fringe-rate filtered, we have substantially increased our integration time, resulting in fewer independent samples (from $\sim700$ samples to $8$ samples). A random draw of $700$ samples from this dataset therefore has many repeated samples, and the variance between hundreds ($N_{boot}$) of these random samples is smaller than the true underlying variance of the data.

Given our new understanding of the sensitivity of bootstraps to the number of elements sampled, we have removed the second step entirely and now simply bootstrap over baselines with the ``single replacement" strategy described in Section \ref{sec:ErrorOverview}.

The single replacement strategy also provides a higher sensitivity estimate as only at most one baseline is repeated in each group. In \citet{ali_et_al2015}, baselines were chosen for groups randomly with replacement. In doing so, $\sim$$3$-$4$ baselines per group were repeated (Figure \ref{fig:toy_error2} shows the fraction of independent samples to be $\sim$$55\%$ for $10$ total independent samples) which sacrificed some sensitivity.

Power spectrum estimates with these bootstrapping changes for PAPER-64 fringe-rate filtered data are shown in Figure \ref{fig:data_errors}. The estimates are unweighted in order to disentangle the effects of bootstrapping from signal loss. As shown in the figure, when the baseline average includes many repeated samples the error is over-estimated, while when only one sample in the same average is allowed the possibility of being a repeat, the error on the noise-dominated higher $k$ modes converges to the theoretical noise level.  However, if more elements are drawn for each bootstrap than the number of independent samples (by drawing many times along the time axis), repeated values begin to crop up and the apparent variation between bootstraps drops, resulting in limits below the predicted noise level.

@@

\subsection{PAPER-64: Bias}
\label{sec:Bias}

@@

Although current $21$ cm published power spectrum upper limits lie a $\sim$couple orders of magnitude above predicted EoR levels, ongoing analyses of deeper sensitivity datasets from PAPER, MWA, and LOFAR, as well as next generation instruments like HERA, expect to continue to push towards EoR sensitivities. As the field progresses towards a coveted detection, we have shown that it is crucial for future analyses to have a rigorous understanding of signal loss in an analysis pipeline, be able to accurately and robustly calculate both power spectrum and theoretical errors, and consistently undergo a comprehensive set of jackknive and null tests.

In particular, in this paper we have investigated the subtleties and tradeoffs of common $21$ cm power spectrum techniques on signal loss, error estimation, and bias, which can be summarized as follows:

\begin{itemize}
\item Substantial signal loss can result when weighting data using empirically estimated covariances. Loss of the $21$ cm signal is especially significant if more eigenmodes are down-weighted than exist in the data (having an under-constrained covariance matrix). Hence, there exists a trade-off between sensitivity driven time-averaging techniques such as fringe-rate filtering and signal loss. 
\item Signal injection and recovery simulations can be used to quantify signal loss. However, a signal-only simulation can underestimate loss by failing to account for spurious correlations between the data and signal. 
\item Errors that are estimated via bootstrapping can be underestimated if samples in the dataset are correlated. If the number of independent samples in a dataset is well-determined, bootstrapping is a simple and accurate way of estimating errors.
\item An EoR detection must pass all null tests. Similarly, performing jackknife tests along multiple axes of a dataset is necessary for confidence in a true EoR detection, and can also be used to tease out systematics.
\end{itemize}
